Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes and ions

### 2019-05-07 Manuscript/manuscript.tex
Small edits in intro

%%SAMULI: new suggestion for the preliminary title
\title[] 
  {Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane}

@@

The phophatidylserine (PS) lipids are most common negatively charged lipids in eukaryotic membranes
and important signaling molecules \cite{lemmon08,leventis10,li14}.
They interact with signaling proteins \cite{leventis10},
regulate surface charge and protein localization \cite{yeung08}, 
induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{??}.
Because these functions are also often regulated by biological cations \cite{leventis10},
the detailed interactions between PS lipids and biological cations, like calcium, are under great interest.
\todo{Pavel can probably improve and complement this}

%Various interaction schemes have been proposed for the negatively charged
%lipids with ions and other lipids. 
Spectroscopic experiments give detailed information about the
interactions between ions and PS lipids, but the data is often indirect and difficult to
interpret \cite{hauser77,kurland79,eisenberg79,hauser83,dluhy83,hauser85,feigenson86,mattai89,roux90,roux91}.
Some studies suggest that the
binding constant of ions to negatively charged lipids is similar to that of zwitterionic lipids,
and the binding affinity is increased only due to the increased cation
concentration in the vicinity of the membrane \cite{seelig90,sinn06}.
On the other hand, calcium forms dehydrated complexes with PS headgroups
which cause phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
Different theories are difficult to evaluate experimentally,
especially with physiological ion concentrations which are typically quite low.
More recently, classical molecular dynamics (MD) simulations
have been used to support the interpretation of spectroscopic experiments,
but the results strongly depend on the force field parameters \cite{boettcher11,kucerka14,melcrova16,hallock18,valentine18}.
In simulations with the CHARMM36 force field \cite{klauda10,venable13} using the NBfix parameters for calcium \cite{kim16}, 
calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
However, the same force field without the NBfix parameters, gives a significant binding affinity also to the phosphate region \cite{hallock18}.
On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
suggest a significant calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.

The NMRlipids project \url{nmrlipids.blogspot.fi} has recently demonstrated
that such controversies can be resolved by using the headgroup order parameters
of phosphatidylcholine (PC) lipids \cite{catte16, NMRlipidsIV}, which can
be related to cation binding affinity to lipid bilayers using the electrometer
concept \cite{akutsu81,altenbach84,seelig87}. The main advantage of the approach
is the direct comparison between experimental and calculated order parameters,
which reduced the ambiguity arising from the interpretation of the data.
Unfortunately, none of the readily available force fields was sufficiently accurate to correctly reproduce the
cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
mixtures with negatively charge PS lipids \cite{NMRlipidsIV}.
In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
to zwitterionic 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC) lipid bilayer
by implicitly including electronic polarizability using the electronic continuum
correction (ECC) \cite{leontyev09}. The good agreement between the resulting ECC-POPC
model and experiments enable detailed interpretation of calcium binding details to POPC lipid bilayer.

@@

Here, we extend the ECC approach also to negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
We also complement the available experimental data for the the force field quality
evaluation by measuring the acyl chain C--H bond order parameters of POPS in bilayer using natural abundance $^{13}$C NMR.
In addition to the acyl chain order parameters measured here, the quality of the newly developed ECC-POPS force field parameters is
evaluated using previously published C--H bond order parameters of glycerol backbone and headgroup in various ionic
conditions and lipid molar fractions \cite{roux90,NMRlipidsIV}, as well as X-ray scattering form factors\cite{kucerka14}.
The overall improvement of the force field accuracy upon applying ECC
%not only describes the interactions of PS with ions more accurately, 
%but also improves the structure of the PS head group 
%and its interactions with PC in mixed lipid bilayers. 
%The results
paves the way to more realistic MD simulations 
of both neutral and negatively charged lipid bilayers 
for a wide range of applications.

>>>>

%%SAMULI: new suggestion for the preliminary title
\title[] 
  { Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes and ions }

@@

Phosphatidylserine (PS) lipids are the most common negatively charged lipids in eukaryotic membranes
and important signaling molecules \cite{lemmon08,leventis10,li14}.
They interact with signaling proteins \cite{leventis10},
regulate surface charge and protein localization \cite{yeung08}, 
induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{??}.
Because these functions are also often regulated by biological cations \cite{leventis10},
the detailed interactions between PS lipids and biological cations, like calcium, are under great interest.
\todo{Pavel can probably improve and complement this}

%Various interaction schemes have been proposed for the negatively charged
%lipids with ions and other lipids. 
Spectroscopic experiments give detailed information about the
interactions between ions and PS lipids, but the data is often indirect and difficult to
interpret \cite{hauser77,kurland79,eisenberg79,hauser83,dluhy83,hauser85,feigenson86,mattai89,roux90,roux91}.
Some studies suggest that the
binding constant of ions to negatively charged lipids is similar to that of zwitterionic lipids,
and the binding affinity is increased only due to the increased cation
concentration in the vicinity of the membrane \cite{seelig90,sinn06}.
On the other hand, calcium forms dehydrated complexes with PS headgroups
which cause phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
Theories about the interaction between calcium and PS 
are difficult to evaluate experimentally,
especially at physiological ion concentrations,
which are typically too low to yield measurable effects.

More recently, classical molecular dynamics (MD) simulations
have been used to support the interpretation of spectroscopic experiments,
but the results strongly depend on the force field parameters \cite{boettcher11,kucerka14,melcrova16,hallock18,valentine18}.
For instance, the moieties of PS lipids that interact with calcium cations vary greatly.
In simulations with the CHARMM36 force field \cite{klauda10,venable13} using the NBfix parameters for calcium \cite{kim16}, 
calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
However, the same force field without the NBfix parameters, gives a significant binding affinity also to the phosphate region \cite{hallock18}.
On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
suggest a significant calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.

The NMRlipids project \url{nmrlipids.blogspot.fi} has recently demonstrated
that such controversies can be resolved by using the headgroup order parameters
of phosphatidylcholine (PC) lipids \cite{catte16, NMRlipidsIV}, which can
be related to cation binding affinity to lipid bilayers using the electrometer
concept \cite{akutsu81,altenbach84,seelig87}. The main advantage of the approach
is the direct comparison between experimental and calculated order parameters,
which reduced the ambiguity arising from the interpretation of the data.
Unfortunately, none of the readily available force fields was sufficiently accurate to correctly reproduce the
cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
mixtures with negatively charged PS lipids \cite{NMRlipidsIV}.
In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
to zwitterionic 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC) lipid bilayer
by implicitly including electronic polarizability using the electronic continuum
correction (ECC) \cite{leontyev09}. The good agreement between the resulting ECC-POPC
model and experiments enable detailed interpretation of calcium binding details to POPC lipid bilayer.

@@

Here, we extend the ECC approach also to the negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
We also complement the available experimental data for the the force field quality
evaluation by measuring the acyl chain C--H bond order parameters of POPS in bilayer using natural abundance $^{13}$C NMR.
In addition to the acyl chain order parameters measured here, the quality of the newly developed ECC-POPS force field parameters is
evaluated using previously published C--H bond order parameters of glycerol backbone and headgroup in various ionic
conditions and lipid molar fractions \cite{roux90,NMRlipidsIV}, as well as X-ray scattering form factors\cite{kucerka14}.
The overall improvement of the force field accuracy upon applying ECC
%not only describes the interactions of PS with ions more accurately, 
%but also improves the structure of the PS head group 
%and its interactions with PC in mixed lipid bilayers. 
%The results
paves the way to more realistic MD simulations 
of both neutral and negatively charged lipid bilayers 
for a wide range of applications.
